### 2017-04-25 paper/paper.tex
60char

\documentclass{paper}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
\Large{\mytitle}
\end{center}

\tableofcontents

\begin{flushleft}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Notations}

For each $n \in \Zplus$:
\begin{enumerate}
\item let set $\sp{n} = \FZnsTwo =
  \set{f: \Zn \times \Zn \to \Ftwo}$
\item let function $\lf{n}: \sp{n+2} \to \sp{n}$
\item let proposition $\cs{n}$ be: $\lf{n}$ is surjective
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}

\begin{remark}
$precell \to cell$:
\begin{itemize}
\item $0 \to 1$ $$\binom{8}{3}$$
\item $0 \to 0$ $$2^8 - \binom{8}{3}$$
\item $1 \to 1$ $$\binom{8}{2} + \binom{8}{3}$$
\item $1 \to 0$ $$2^8 - \binom{8}{2} - \binom{8}{3}$$
\end{itemize}
\end{remark}

\begin{remark}
For each $n \in \set{1, 2, 3, 4}$, $\cs{n}$ is true.
\end{remark}

\begin{remark}
For each $n \in \set{5, 6}$, $\cs{n}$ is
\emph{probably} true.
\end{remark}

\begin{conjecture}
For each $n \in \Zplus$, $\cs{n}$ is true.
\end{conjecture}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{flushleft}

\end{document}