### 2017-04-26 paper/paper.tex
rearranged sections

\documentclass[letterpaper, 14pt]{extarticle}
% 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, 20pt

\usepackage[scale=0.85]{geometry}

\usepackage{paper}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
\Large{\mytitle}
\end{center}

\tableofcontents

\begin{flushleft}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Combinatorics}

\begin{remark}
case count: precell state $\to$ cell state:
\begin{itemize}
\item $0 \to 1$ $$\binom{8}{3}$$
\item $0 \to 0$ $$2^8 - \binom{8}{3}$$
\item $1 \to 1$ $$\binom{8}{2} + \binom{8}{3}$$
\item $1 \to 0$ $$2^8 - \binom{8}{2} - \binom{8}{3}$$
\end{itemize}
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Game}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Cell}

\begin{definition}
$f(x)$
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Grid}

\begin{definition}
For each $n \in \Zplus$:
\begin{enumerate}
\item let set $\sp{n} = \FZnsTwo =
  \set{f: \Zn \times \Zn \to \Ftwo}$
\item let function $\lf{n}: \sp{n+2} \to \sp{n}$
\item let proposition $\cs{n}$ be: $\lf{n}$ is surjective
\end{enumerate}
\end{definition}

\begin{remark}
$\lf{n}$ sizes:
\begin{itemize}
\item domain: $2^{(n+2)^2}$
\item codomain: $2^{n^2}$
\item average fiber: $2^{4(n+1)}$
\end{itemize}
\end{remark}

\begin{remark}
For each $n \in \set{1, 2, 3, 4}$, $\cs{n}$ is true.
\end{remark}

\begin{remark}
For each $n \in \set{5, 6}$, $\cs{n}$ is
\emph{probably} true.
\end{remark}

\begin{conjecture}
For each $n \in \Zplus$, $\cs{n}$ is true.
\end{conjecture}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{flushleft}

\end{document}