### 2017-05-12 paper/paper.tex
Dr. Monico checked 5x5 in 5 days

\documentclass[letterpaper, 17pt]{extarticle}
% 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, 20pt

\usepackage[scale=0.9]{geometry}

\usepackage{paper}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
\Large{\mytitle}
\end{center}

\tableofcontents

\begin{flushleft}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Combinatorics}

\begin{remark}
case count: precell state $\to$ cell state:
\begin{itemize}
\item $0 \to 1$ $$\binom{8}{3}$$
\item $0 \to 0$ $$2^8 - \binom{8}{3}$$
\item $1 \to 1$ $$\binom{8}{2} + \binom{8}{3}$$
\item $1 \to 0$ $$2^8 - \binom{8}{2} - \binom{8}{3}$$
\end{itemize}
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Game}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Cell}

\begin{definition}
Life polynomial:
\begin{itemize}
\item Let cell $\cell \in \Ft$ and
  neighbors $\neighbors \in \Ft$.
\item Let polynomial $\lp \in \pr$.
\item Let polynomial function
  $\pf: \Ft \times \Fte \to \Ft$.
\end{itemize}
\end{definition}

\begin{remark}
\begin{dmath*}
\lp = (\cell + 1) *
\sum_{i \ne j \ne k, k \ne i}\parentheses
{\cell_i \cell_j \cell_k
\prod_{m \ne i, j, k}(\cell_m + 1)} + \\
\cell * \parentheses
{\sum_{i \ne j}\parentheses
{\cell_i \cell_j \prod_{m \ne i, j}(\cell_m + 1)} + \\
\sum_{i \ne j \ne k, k \ne i}\parentheses
{\cell_i \cell_j \cell_k
\prod_{m \ne i, j, k}(\cell_m + 1)}}
\end{dmath*}
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Grid}

\begin{definition}
For each $n \in \Zplus$:
\begin{itemize}
\item Let set $\sp{n} = \FtZns =
  \set{grid: \Zn \times \Zn \to \Ft}$.
\item Let function $\lf{n}: \sp{n+2} \to \sp{n}$.
\end{itemize}
\end{definition}

\begin{remark}
$\lf{n}$ sizes:
\begin{itemize}
\item domain: $2^{(n+2)^2}$
\item codomain: $2^{n^2}$
\item average fiber: $2^{4(n+1)}$
\end{itemize}
\end{remark}

\begin{definition}
Let proposition $\cs{n}$ be: $\lf{n}$ is surjective.
\end{definition}

\begin{remark}
For each $n \in \set{1, 2, 3, 4}$, $\cs{n}$ is true.
\end{remark}

\begin{remark}
Dr. Monico: $\cs5$ is true (5 days).
\end{remark}

\begin{remark}
Vu: $\cs6$ is \emph{likely} true \\
(1\% complete, 10 days elapsed, 3 years left).
\end{remark}

\begin{conjecture}
For each $n \in \Zplus$, $\cs{n}$ is true.
\end{conjecture}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{flushleft}

\end{document}