### 2017-05-15 paper/paper.tex
style

\documentclass[letterpaper, 14pt]{extarticle}
% 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, 20pt

\usepackage[scale=0.9]{geometry}

\usepackage{paper}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
\Large{\mytitle}
\end{center}

\tableofcontents

\begin{flushleft}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Counting}

\begin{remark}
Case counts (precell state $\to$ cell state):
\begin{itemize}
\item $0 \to 1$ $$\binom{8}{3}$$
\item $0 \to 0$ $$2^8 - \binom{8}{3}$$
\item $1 \to 1$ $$\binom{8}{2} + \binom{8}{3}$$
\item $1 \to 0$ $$2^8 - \binom{8}{2} - \binom{8}{3}$$
\end{itemize}
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Functions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Cell Function}

\begin{definition}
Polynomial function:
\begin{itemize}
\item Let cell $\cell \in \Ft$ and
  neighbors $\neighbors \in \Ft$.
\item Let polynomial $\lp \in \pr$.
\item Let \textdef{cell function}
  $\pf: \Ft \times \Fte \to \Ft$.
\end{itemize}
\end{definition}

\begin{remark}
Explicit form:
\begin{dmath*}
\lp = (\cell + 1) *
\sum_{i \ne j \ne k, k \ne i}\parentheses
{\cell_i \cell_j \cell_k
\prod_{m \ne i, j, k}(\cell_m + 1)} + \\
\cell * \parentheses
{\sum_{i \ne j}\parentheses
{\cell_i \cell_j \prod_{m \ne i, j}(\cell_m + 1)} + \\
\sum_{i \ne j \ne k, k \ne i}\parentheses
{\cell_i \cell_j \cell_k
\prod_{m \ne i, j, k}(\cell_m + 1)}}
\end{dmath*}
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Grid Function}

\begin{definition}
For each $n \in \Zplus$:
\begin{itemize}
\item Let set $\sp{n} = \FtZns =
  \set{grid: \Zn \times \Zn \to \Ft}$.
\item Let \textdef{grid function} $\lf{n}: \sp{n+2} \to \sp{n}$.
\end{itemize}
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Cardinalities}

\begin{remark}
Cardinalities related to $\lf{n}$:
\begin{itemize}
\item domain: $D_n = 2^{(n+2)^2}$
\item codomain: $C_n = 2^{n^2}$
\item image: $I_n \le C_n$
\item proportion: $P_n = I_n / C_n \le 1$
\item average fiber: $D_n / I_n \ge D_n / C_n = 2^{4(n+1)}$
\end{itemize}
\end{remark}

\begin{remark}[joining small grids into a big grid]
Let $k \in \Zplus$. Then:
$$P_{k n} = \frac{I_{k n}}{C_{k n}}
= \frac{I_{k n}}{(C_n)^{k^2}}
\le \frac{{(I_n)}^{k^2}}{(C_n)^{k^2}}
= \parentheses{\frac{I_n}{C_n}}^{k^2}
= (P_n)^{k^2}$$
If $P_n < 1$ for some $n$,
then $\lim_{k \to \infty} P_{kn} = 0$.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Surjectivity}

\begin{definition}
Let proposition $\cs{n}$ be: $\lf{n}$ is surjective.
\end{definition}

\begin{remark}
For each $n \in \set{1, 2, 3, 4}$, $\cs{n}$ is true.
\end{remark}

\begin{remark}
\cm's code: $\cs5$ is true (5-day runtime).
\end{remark}

\begin{remark}
Vu's code: $\cs6$ is \emph{likely} true \\
(1\% complete, 10 days elapsed, 3 years left).
\end{remark}

\begin{conjecture}
For each $n \in \Zplus$, $\cs{n}$ is true.
\end{conjecture}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{flushleft}

\end{document}